---
title: "Romero_Lab2"
author: "Guillermo Romero"
date: "2023-01-18"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
# Source from lab 1
sourceDir <- "/Users/guillermoromero/Documents/MEDS/eds232_ml/Romero_Lab1.Rmd"
library(knitr)
source(knitr::purl(sourceDir, quiet=TRUE))
```

Today we will be continuing the pumpkin case study from last week. We
will be using the data that you cleaned and split last time
(pumpkins_train) and will be comparing our results today to those you
have already obtained, so open and run your Lab 1 .Rmd as a first step
so those objects are available in your Environment (unless you created
an R Project last time, in which case, kudos to you!).

Once you have done that, we'll start today's lab by specifying a recipe
for a polynomial model. First we specify a recipe that identifies our
variables and data, converts package to a numerical form, and then add a
polynomial effect with step_poly()

```{r}
# Specify a recipe
poly_pumpkins_recipe <-
  recipe(price ~ package, data = pumpkins_train) %>%
  step_integer(all_predictors(), zero_based = TRUE) %>% 
  step_poly(all_predictors(), degree = 4)
```

How did that work? Choose another value for degree if you need to. Later
we will learn about model tuning that will let us do things like find
the optimal value for degree. For now, we'd like to have a flexible
model, so find the highest value for degree that is consistent with our
data.

Polynomial regression is still linear regression, so our model
specification looks similar to before.

```{r}
# Create a model specification called poly_spec
poly_spec <- linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")
```

Question 1: Now take the recipe and model specification that just
created and bundle them into a workflow called poly_df.

```{r}
# Bundle recipe and model spec into a workflow
poly_wf <- workflow() %>% 
  add_recipe(poly_pumpkins_recipe) %>% 
  add_model(poly_spec)
```

Question 2: fit a model to the pumpkins_train data using your workflow
and assign it to poly_wf_fit

```{r}
# Create a model
poly_wf_fit <- fit(poly_wf, data = pumpkins_train)
```

```{r}
# Print learned model coefficients
poly_wf_fit
```

```{r}
# Make price predictions on test data
poly_results <- poly_wf_fit %>% predict(new_data = pumpkins_test) %>% 
  bind_cols(pumpkins_test %>% select(c(package, price))) %>% 
  relocate(.pred, .after = last_col())

# Print the results
poly_results %>% 
  slice_head(n = 10)
```

Now let's evaluate how the model performed on the test_set using
yardstick::metrics().

```{r}
metrics(data = poly_results, truth = price, estimate = .pred)
```

Question 3: How do the performance metrics differ between the linear
model from last week and the polynomial model we fit today? Which model
performs better on predicting the price of different packages of
pumpkins?

***The performance metrics are all much lower for all the polynomial fit
when compared to the linear model from last week. The root mean squared
was 7.23, the average price by which the predictions are wrong*** ***and
lower at 3.27 this week. The polynomial fit model performs better on
predicting the price of different packages of pumpkins.***

Let's visualize our model results. First prep the results by binding the
encoded package variable to them.

```{r}
# Bind encoded package column to the results
poly_results <- poly_results %>% 
  bind_cols(package_encode %>% 
              rename(package_integer = package)) %>% 
  relocate(package_integer, .after = package)


# Print new results data frame
poly_results %>% 
  slice_head(n = 5)
```

OK, now let's take a look!

Question 4: Create a scatter plot that takes the poly_results and plots
package vs. price. Then draw a line showing our model's predicted values
(.pred). Hint: you'll need separate geoms for the data points and the
prediction line.

```{r}
# Make a scatter plot
poly_results %>%
  ggplot(mapping = aes(x = package_integer, y = price)) +
   geom_point(size = 1.6) +
   # Overlay a line of best fit
   geom_line(aes(y = .pred), color = "orange", size = 1.2) +
   xlab("package")
```

You can see that a curved line fits your data much better.

Question 5: Now make a smoother line by using geom_smooth instead of
geom_line and passing it a polynomial formula like this:
geom_smooth(method = lm, formula = y \~ poly(x, degree = 3), color =
"midnightblue", size = 1.2, se = FALSE)

```{r}
# Make a smoother scatter plot 
poly_results %>%
  ggplot(mapping = aes(x = package_integer, y = price)) +
   geom_point(size = 1.6) +
  geom_smooth(method = lm, formula = y ~ poly(x, degree = 3), color = "midnightblue", size = 1.2, se = FALSE)
```

OK, now it's your turn to go through the process one more time.

Additional assignment components :

6.  Choose a new predictor variable (anything not involving package
    type) in this dataset. 7. Determine its correlation with the outcome
    variable (price). (Remember we calculated a correlation matrix last
    week)

```{r}
cor(baked_pumpkins$day, baked_pumpkins$price)
```

8.  Create and test a model for your new predictor: - Create a recipe -
    Build a model specification (linear or polynomial) - Bundle the
    recipe and model specification into a workflow - Create a model by
    fitting the workflow - Evaluate model performance on the test data -
    Create a visualization of model performance

```{r}
#Create a recipe
poly_recipe_day <-
  recipe(price ~ day, data = pumpkins_train) %>%
  step_integer(all_predictors(), zero_based = TRUE) %>% 
  step_poly(all_predictors(), degree = 4)
```

```{r}
# Create a model specification 
poly_spec_day <- linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")
```

```{r}
# Bundle recipe and model spec into a workflow
poly_wf_day <- workflow() %>% 
  add_recipe(poly_recipe_day) %>% 
  add_model(poly_spec_day)
poly_wf_day
```

```{r}
# Create a model by fitting the workflow
poly_fit_day <- fit(poly_wf_day, data = pumpkins_train)
```

```{r}
# Make price predictions on test data
poly_results_day <- poly_fit_day %>% predict(new_data = pumpkins_test) %>% 
  bind_cols(pumpkins_test %>% select(c(day, price))) %>% 
  relocate(.pred, .after = last_col())

# Print new results data frame
poly_results_day %>% 
  slice_head(n = 5)
```

```{r}
#Evaluate model performance on the test data
metrics(data = poly_results_day, truth = price, estimate = .pred)
```

***The model performance metrics are all worse, the rmse using the
predictor 'variety' with a polynomial fit is higher than that of package
type 9.53 vs. 3.27 and is worse than the linear regression from last
week.***

```{r}
# Create a visualization of model performance

poly_results_day %>%
  ggplot(mapping = aes(x = day, y = price)) +
   geom_point(size = 1.6, colour = 'black', alpha = 0.6,
              stroke = .8,fill = 'dodgerblue',
              shape = 21) +
  geom_smooth(method = lm, formula = y ~ poly(x, degree = 4), color = "midnightblue", size = 1.2, se = FALSE) +
  theme_classic()
```

Lab 2 due 1/24 at 11:59 PM
